Rat E2 Project
========================================================
Brain gene expression
-----------------------------------------------
First we can clean up the data
```{r}
##############functions#################

#report columns that are missing too much data

removeCol<-function(mat){
  badCol<-vector()
  for(col in 1:ncol(mat)){
    if(all(is.na(mat[,col]))) (badCol<-c(badCol,col)) 
  }
  return(badCol)
}
removeRow<-function(mat){
  badRow<-vector()
  for(row in 1:nrow(mat)){
    if(all(is.na(mat[row,]))) (badRow<-c(badRow,row)) 
  }
  return(badRow)
}

fixNames<-function(names,brain_area){
 gene<-vector()
 for(probe in names){
   gsplt<-unlist(strsplit(probe,split=".",fixed=TRUE))
   gsplt<-paste(brain_area,gsplt,sep=".")
   gene<-c(gene,gsplt[1])
 }
 return(gene)
}

fixNames2<-function(names){
 gene<-vector()
 for(probe in names){
   gsplt<-unlist(strsplit(probe,split=".",fixed=TRUE))
   gene<-c(gene,gsplt[1])
 }
 return(gene)
}


fix.cor.mat<-function(corgram){
  fix<-diag(8)
  fix[8,1:7]<-NA
  fix[2:8,1]<-NA
  for(row in 1:nrow(corgram)){
    for(col in 1:ncol(corgram)){
      rowu<-as.numeric(rownames(corgram))[row]
      colu<-as.numeric(colnames(corgram))[col]
      if(rowu!=colu) (fix[rowu,colu]<-corgram[row,col])
    }
  }
  return(fix)
}

makeCorGram_withinGroup<-function(poa,poa.mat){ # should refactor this to use generic names
require(reshape2)
across.groups.poa<-as.data.frame(t(combn(1:8,2)))
corrl<-vector()
for(i in 1:nrow(across.groups.poa)){
  poa.1<-cov(poa.mat[poa$Group.ID==across.groups.poa[i,1],],use="pairwise.complete.obs")
  poa.2<-cov(poa.mat[poa$Group.ID==across.groups.poa[i,2],],use="pairwise.complete.obs")
  corrl<-c(corrl,cor(as.numeric(diag.remove(poa.1)),as.numeric(diag.remove(poa.2)),use="pairwise.complete.obs"))  
}
across.groups.poa$correlation<-corrl

poa.across.groups.corgram<-dcast(across.groups.poa,V1~V2)
rownames(poa.across.groups.corgram)<-poa.across.groups.corgram[,1]
poa.across.groups.corgram<-poa.across.groups.corgram[,-1]
poa.fix.corgram<-fix.cor.mat(poa.across.groups.corgram)
return(poa.fix.corgram)
}

makeCorGram_AcrossBrain<-function(poa,poa.mat,arc,arc.mat.small){
require(reshape2)
across.brain<-as.matrix(data.frame(group.arc=1:8,group.poa=1:8))
across.brain<-rbind(across.brain,t(combn(1:8,2)))
corrl<-vector()
for(i in 1:nrow(across.brain)){
  arc.1<-cov(arc.mat.small[arc$Group.ID==across.brain[i,1],],use="pairwise.complete.obs")
  poa.1<-cov(poa.mat[poa$Group.ID==across.brain[i,2],],use="pairwise.complete.obs")
  corrl<-c(corrl,cor(as.numeric(diag.remove(arc.1)),as.numeric(diag.remove(poa.1)),use="pairwise.complete.obs"))  
}
across.brain<-as.data.frame(across.brain)
across.brain$correlation<-corrl

across.brain.corgram<-dcast(across.brain,group.arc~group.poa)
rownames(across.brain.corgram)<-across.brain.corgram[,1]
across.brain.corgram<-across.brain.corgram[,-1]
return(across.brain.corgram)
}

#########################################

library(sna)
arc<-read.csv("ARC TLDA for stats 7-20-13.csv") # load the data
arc<-arc[1:82,]
poa<-read.csv("mPOA TLDA for stats 7-20-13.csv")

names(arc)[18]
names(arc)[64] #bounds are 18 to 64

arc.mat<-arc[,18:64] # column 11 is missing data.
which(grepl("Mtnr1a",names(arc.mat)))
arc.mat[,30]
which(arc.mat=="#VALUE!",arr.ind=TRUE)
arc.mat[74,30]<-NA
arc.mat[,30]<-as.numeric(as.character(arc.mat[,30]))

arc.mat[,colnames(arc.mat)=="Dbh.Rn00565819_m1"]
removeCol(arc.mat)

#arc.mat.small<-(arc.mat[,c(-11,-31,-33)]) # this is to make it match witht he POA data
#arc.mat<-(arc.mat[,-11])
#arc.mat<-as.matrix(arc.mat)

poa.mat<-(poa[,17:63])
str(poa.mat) #Dbh, Mtnr1b, Npvf
which(grepl("Dbh",names(poa.mat))) #11
which(grepl("Mtnr1b",names(poa.mat))) #31
which(grepl("Npvf",names(poa.mat))) #33
removeCol(poa.mat)

poa.mat<-as.matrix(poa.mat[,c(-11,-31,-33)])
arc.mat.small<-as.matrix(arc.mat[,c(-11,-31,-33)])
arc.mat<-as.matrix(arc.mat[,-11])

unique(colnames(poa.mat)==colnames(arc.mat.small))

#colnames(arc.mat.small)<-fixNames(colnames(arc.mat.small),"arc")
#colnames(poa.mat)<-fixNames(colnames(poa.mat),"poa")

rownames(arc.mat.small)<-fixNames(arc$Sample.Name[1:82],"arc")
rownames(poa.mat)<-fixNames(poa$Sample.Name,"poa")

fullMat<-rbind(arc.mat.small,poa.mat)
dim(arc.mat.small)
dim(poa.mat)
dim(fullMat)

```

```{r}
badRows<-removeRow(fullMat)
fullMat<-fullMat[-badRows,]
#imputing missing data
library(imputation)
fullMat<-kNNImpute(fullMat,3)$x
```

```{r fig.height=20, fig.width=20}
library(NeatMap)
make.heatmap1(as.matrix(scale(fullMat)),row.method="nMDS",column.method="nMDS",row.labels=rownames(fullMat),column.labels=colnames(fullMat))
```

```{r fig.height=20, fig.width=20}
rownames(fullMat)<-c(arc$Group.code[1:82][-badRows],poa$Group.code[-badRows])
make.heatmap1(as.matrix(scale(fullMat)),row.method="nMDS",column.method="nMDS",row.labels=rownames(fullMat),column.labels=colnames(fullMat))
```

## POA and Arcuate cluster together much more than treatment cluster together


## Covariance networks
```{r}
poa.cov<-cov(poa.mat,use="pairwise.complete.obs")
arc.cov<-cov(arc.mat.small,use="pairwise.complete.obs")

graphCorrelation<-cor(as.numeric(diag.remove(poa.cov)),as.numeric(diag.remove(arc.cov)),use="pairwise.complete.obs")
paste("Overall corrleation of the covariance networks between the brain regions is",graphCorrelation,sep=" ")
```

Groups are as follows:
1. YV3
2. YE3
3. MV3
4. ME3
5. MV6
6. ME6
7. MV3E3
8. ME3V3

## Similarity of covariance networks within POA, across treatment groups
```{r fig.height=6, fig.width=6}
poa.fix.corgram<-makeCorGram_withinGroup(poa,poa.mat)
print(levelplot(as.matrix(poa.fix.corgram),xlab="poa.treatment.group",ylab="poa.treatment.group",at=do.breaks(c(-1.01,1.01),101),scales=list(x=list(rot=90)),colorkey=list(space="top"),col.regions=colorRampPalette(c("gold","gray85","purple2"))))
```

## Similarity of covariance networks within ARC, across treatment groups
```{r fig.height=6, fig.width=6}
arc.fix.corgram<-makeCorGram_withinGroup(arc,arc.mat.small)
print(levelplot(as.matrix(arc.fix.corgram),xlab="arc.treatment.groups",ylab="arc.treatment.groups",at=do.breaks(c(-1.01,1.01),101),scales=list(x=list(rot=90)),colorkey=list(space="top"),col.regions=colorRampPalette(c("gold","gray85","purple2"))))
```

## Similarity of covariance networks across Brain areas and treatment groups
```{r fig.height=6, fig.width=6}
across.brain<-makeCorGram_AcrossBrain(poa,poa.mat,arc,arc.mat.small)
print(levelplot(as.matrix(across.brain),xlab="arc.treatment.groups",ylab="poa.treatment.groups",at=do.breaks(c(-1.01,1.01),101),scales=list(x=list(rot=90)),colorkey=list(space="top"),col.regions=colorRampPalette(c("gold","gray85","purple2"))))
```

## Visualizing Covariance networks

```{r}
library(igraph)
poa.cor<-cor(poa.mat,use="pairwise.complete.obs")
arc.cor<-cor(arc.mat,use="pairwise.complete.obs")

rownames(poa.cor)<-fixNames2(rownames(poa.cor))
colnames(poa.cor)<-fixNames2(colnames(poa.cor))

rownames(arc.cor)<-fixNames2(rownames(arc.cor))
colnames(arc.cor)<-fixNames2(colnames(arc.cor))

library(psych)
poa.cor.test<-corr.test(poa.mat)$p
poa.cor.test[lower.tri(poa.cor.test)]<-poa.cor.test[upper.tri(poa.cor.test)]
badCell<-as.data.frame(which(poa.cor.test>.01,arr.ind=TRUE,useNames=FALSE))
dim(poa.cor)
dim(poa.cor.test)

for(i in 1:length(badCell[,1])){
  poa.cor[badCell[i,1],badCell[i,2]]<-0
}

poa.network<-graph.adjacency(poa.cor,mode="undirected",weighted=TRUE,diag=FALSE)

minC <- rep(-Inf, vcount(poa.network))
maxC <- rep(Inf, vcount(poa.network))
minC[1] <- maxC[1] <- 0
co <- layout.fruchterman.reingold(poa.network, minx=minC, maxx=maxC,
                                  miny=minC, maxy=maxC)
```
```{r fig.height=15,fig.width=15}
plot(poa.network,layout=co,edge.width=5*get.edge.attribute(poa.network,"weight"))
```
```{r}
arc.cor.test<-corr.test(arc.mat)$p
arc.cor.test[lower.tri(arc.cor.test)]<-arc.cor.test[upper.tri(arc.cor.test)]
badCell<-as.data.frame(which(arc.cor.test>.01,arr.ind=TRUE,useNames=FALSE))
dim(arc.cor)
dim(arc.cor.test)

for(i in 1:length(badCell[,1])){
  arc.cor[badCell[i,1],badCell[i,2]]<-0
}

arc.network<-graph.adjacency(arc.cor,mode="undirected",weighted=TRUE,diag=FALSE)

minC <- rep(-Inf, vcount(arc.network))
maxC <- rep(Inf, vcount(arc.network))
minC[1] <- maxC[1] <- 0
co <- layout.fruchterman.reingold(arc.network, minx=minC, maxx=maxC,
                                  miny=minC, maxy=maxC)
```
```{r  fig.height=15,fig.width=15}
plot(arc.network,layout=co,edge.width=5*get.edge.attribute(arc.network,"weight"))
```

```{r}
hist(degree(poa.network))
hist(degree(arc.network))
```


WGCNA
===================================================================

Install first time only:
source("http://bioconductor.org/biocLite.R")
biocLite("impute")
install.packages("WGCNA") 
```{r}
library(WGCNA)
allowWGCNAThreads() # faster w/ multiple threads

#1 fix names.
gene<-vector()
for(probe in colnames(arc.mat)){
  gsplt<-unlist(strsplit(probe,split=".",fixed=TRUE))
  gene<-c(gene,gsplt[1])
}

colnames(arc.mat)<-gene
rownames(arc.mat)<-as.character(arc$Sample.Name)
gsg <- goodSamplesGenes(arc.mat, verbose = 3);

if (!gsg$allOK)
{
  # Optionally, print the gene and sample names that were removed:
  if (sum(!gsg$goodGenes)>0)
    printFlush(paste("Removing genes:", paste(names(arc.mat)[!gsg$goodGenes], collapse = ", ")));
  if (sum(!gsg$goodSamples)>0)
    printFlush(paste("Removing samples:", paste(rownames(arc.mat)[!gsg$goodSamples], collapse = ", ")));
  # Remove the offending genes and samples from the data:
  arc.mat.gsg <- arc.mat[gsg$goodSamples, gsg$goodGenes]
}
#Removing samples: ABV26, ABV75, ABV49, ABV44, ABV45, ABV46, ABV55

sampleTree = flashClust(dist(arc.mat.gsg), method = "average") #Removing ABV86
arc.mat.gsg<-arc.mat.gsg[rownames(arc.mat.gsg)!="ABV86",]
nGenes<-ncol(arc.mat.gsg)
nSamples<-nrow(arc.mat.gsg)

sampNames<-rownames(arc.mat.gsg)
traitRows<- match(sampNames,as.character(arc$Sample.Name))

names(arc)[6:8] <- c("e2", "age", "duration")
arc$e2[61:82] <- as.factor(ifelse(arc$Group.ID[61:82] == 7, yes = "e", 
    no = "v"))
arc$age[which(arc$age == "o")] <- factor("m")
arc$duration <- factor(c(rep(3, length(1:40)), rep(6, length(41:60)), rep(NA,length(61:82))))

arc$duration<-as.factor((as.character(arc$duration)))

datTraits<- data.frame(Treatment=as.numeric(arc$Group.ID[traitRows]),E2=as.numeric(arc$e2[traitRows]),age=as.numeric(arc$age[traitRows]),duration=as.numeric(arc$duration[traitRows]))

sampleTree2 = flashClust(dist(arc.mat.gsg), method = "average")
traitColors = numbers2colors((datTraits), signed = FALSE);

```
```{r fig.height=9,fig.width=12}
plotDendroAndColors(sampleTree2, traitColors,
                    groupLabels = names(datTraits),
                    main = "Sample dendrogram and trait heatmap")
```
```{r}
#powers<- c(1:10, seq(from=12,to=20,by=2))
#sft<-pickSoftThreshold(arc.mat.gsg,powerVector=powers,verbose=5)
#sizeGrWindow(9,5)
#par(mfrow=c(1,2))
#cex1=.9

#plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],xlab="soft threshold (power)",ylab="scale free R2",type="n",main="scale independence")
#text(sft$fitIndices[,1],-sign(sft$fitIndices[,3])*sft$fitIndices[,2],labels=powers,cex=cex1,col="red")
#abline(h=.9,col="red")

#plot(sft$fitIndices[,1],sft$fitIndices[,5],type="n",xlab="power",ylab="mean connectivity")
#text(sft$fitIndices[,1],sft$fitIndices[,5],labels=powers,cex=cex1,col="red")

#I'll go with 4 for now but may test from 4-8

softPower<-5
adjacency=adjacency(arc.mat.gsg,power=softPower)
TOM=TOMsimilarity(adjacency)
dissTOM<- 1-TOM

geneTree=flashClust(as.dist(dissTOM),method="average")
#sizeGrWindow(12,9)
```
```{r fig.height=9,fig.width=12}
plot(geneTree,xlab="",sub="",main="Gene clustering on TOM-based dissimilarity",labels=FALSE,hang=.04)
```
```{r}
minModuleSize=0
dynamicMods<- cutreeDynamic(dendro=geneTree,distM=dissTOM,deepSplit=2,pamRespectsDendro=FALSE,minClusterSize=minModuleSize)
#table(dynamicMods)

dynamicColors=labels2colors(dynamicMods)
table(dynamicColors)
```
```{r fig.height=9, fig.width=12}
plotDendroAndColors(geneTree,dynamicColors,"Dynamic Tree Cut",dendroLabels=FALSE,hang=.03,addGuide=TRUE,guideHang=.05,main="Gene dendrogram and module colors")
```
```{r}
MEList<-moduleEigengenes(arc.mat.gsg,colors=dynamicColors)
Modules<-data.frame(gene=colnames(arc.mat.gsg),color=MEList$validColors)
MEs<-MEList$eigengenes

MEs$treatment<-arc$Group.code[match(rownames(arc.mat.gsg),arc$Sample.Name)]
blue<-aov(MEblue~treatment,data=MEs)
summary(blue)
```
```{r fig.height=3.5,fig.width=5}
plot(MEblue~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
brown<-aov(MEbrown~treatment,data=MEs)
summary(brown)
plot(MEbrown~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
green<-aov(MEgreen~treatment,data=MEs)
summary(green)
plot(MEgreen~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
names(MEs)
grey<-aov(MEgrey~treatment,data=MEs)
summary(grey)
```
```{r fig.height=3.5,fig.width=5}
red<-aov(MEred~treatment,data=MEs)
summary(red)
plot(MEred~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
turquoise<-aov(MEturquoise~treatment,data=MEs)
summary(turquoise)
plot(MEturquoise~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
yellow<-aov(MEyellow~treatment,data=MEs)
summary(yellow)
plot(MEyellow~treatment,data=MEs)
```

```{r}
##################################POA###############################
#1 fix names.
#rm(list=ls())
gene<-vector()
for(probe in colnames(poa.mat)){
  gsplt<-unlist(strsplit(probe,split=".",fixed=TRUE))
  gene<-c(gene,gsplt[1])
}

colnames(poa.mat)<-gene
rownames(poa.mat)<-as.character(poa$Sample.Name)
gsg <- goodSamplesGenes(poa.mat, verbose = 3);

if (!gsg$allOK)
{
  # Optionally, print the gene and sample names that were removed:
  if (sum(!gsg$goodGenes)>0)
    printFlush(paste("Removing genes:", paste(names(poa.mat)[!gsg$goodGenes], collapse = ", ")));
  if (sum(!gsg$goodSamples)>0)
    printFlush(paste("Removing samples:", paste(rownames(poa.mat)[!gsg$goodSamples], collapse = ", ")));
  # Remove the offending genes and samples from the data:
  poa.mat.gsg <- poa.mat[gsg$goodSamples, gsg$goodGenes]
}
#Removing samples: ABV26, ABV85, ABV18, ABV49, ABV43, ABV56

sampleTree = flashClust(dist(poa.mat.gsg), method = "average");
#sizeGrWindow(12,9)
```
```{r fig.height=9,fig.width=12}
plot(sampleTree)
```
```{r}
nGenes<-ncol(poa.mat.gsg)
nSamples<-nrow(poa.mat.gsg)

sampNames<-rownames(poa.mat.gsg)
traitRows<- match(sampNames,as.character(poa$Sample.Name))

names(poa)[6:8] <- c("e2", "age", "duration")
poa$e2[61:82] <- as.factor(ifelse(poa$Group.ID[61:82] == 7, yes = "e", 
    no = "v"))
poa$age[which(poa$age == "o")] <- factor("m")
poa$duration <- factor(c(rep(3, length(1:40)), rep(6, length(41:60)), rep(NA,length(61:82))))

poa$duration<-as.factor((as.character(poa$duration)))

datTraits<- data.frame(Treatment=as.numeric(poa$Group.ID[traitRows]),E2=as.numeric(poa$e2[traitRows]),age=as.numeric(poa$age[traitRows]),duration=as.numeric(poa$duration[traitRows]))

sampleTree2 = flashClust(dist(poa.mat.gsg), method = "average")
traitColors = numbers2colors(datTraits, signed = FALSE);
```
```{r fig.height=9,fig.width=12}
plotDendroAndColors(sampleTree2, traitColors,
                    groupLabels = names(datTraits),
                    main = "Sample dendrogram and trait heatmap")
```
```{r}
#powers<- c(1:10, seq(from=12,to=20,by=2))
#sft<-pickSoftThreshold(poa.mat.gsg,powerVector=powers,verbose=5)
#sizeGrWindow(9,5)
#par(mfrow=c(1,2))
#cex1=.9

#plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],xlab="soft threshold (power)",ylab="scale free R2",type="n",main="scale independence")
#text(sft$fitIndices[,1],-sign(sft$fitIndices[,3])*sft$fitIndices[,2],labels=powers,cex=cex1,col="red")
#abline(h=.9,col="red")
#plot(sft$fitIndices[,1],sft$fitIndices[,5],type="n",xlab="power",ylab="mean connectivity")
#text(sft$fitIndices[,1],sft$fitIndices[,5],labels=powers,cex=cex1,col="red")

#I'll go with 4 for now but may test from 4-8

softPower<-5
adjacency=adjacency(poa.mat.gsg,power=softPower)
TOM=TOMsimilarity(adjacency)
dissTOM<- 1-TOM

geneTree=flashClust(as.dist(dissTOM),method="average")
#sizeGrWindow(12,9)
```
```{r fig.height=9,fig.width=12}
plot(geneTree,xlab="",sub="",main="Gene clustering on TOM-based dissimilarity",labels=FALSE,hang=.04)
```
```{r}
minModuleSize=0
dynamicMods<- cutreeDynamic(dendro=geneTree,distM=dissTOM,deepSplit=2,pamRespectsDendro=FALSE,minClusterSize=minModuleSize)
table(dynamicMods)

dynamicColors=labels2colors(dynamicMods)
table(dynamicColors)
```
```{r fig.height=9,fig.width=12}
plotDendroAndColors(geneTree,dynamicColors,"Dynamic Tree Cut",dendroLabels=FALSE,hang=.03,addGuide=TRUE,guideHang=.05,main="Gene dendrogram and module colors")
```
```{r}
MEList<-moduleEigengenes(poa.mat.gsg,colors=dynamicColors)
Modules<-data.frame(gene=colnames(poa.mat.gsg),color=MEList$validColors)
MEs<-MEList$eigengenes

MEs$treatment<-poa$Group[match(rownames(poa.mat.gsg),poa$Sample.Name)]

names(MEs)
```
```{r fig.height=3.5,fig.width=5}
blue<-aov(MEblue~treatment,data=MEs)
summary(blue)
plot(MEblue~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
brown<-aov(MEbrown~treatment,data=MEs)
summary(brown)
plot(MEbrown~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
green<-aov(MEgreen~treatment,data=MEs)
summary(green)
plot(MEgreen~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
grey<-aov(MEgrey~treatment,data=MEs)
summary(grey)
plot(MEgrey~treatment,data=MEs)
```
```{r}
turquoise<-aov(MEturquoise~treatment,data=MEs)
summary(turquoise)
```
```{r fig.height=3.5,fig.width=5}
yellow<-aov(MEyellow~treatment,data=MEs)
summary(yellow)
plot(MEyellow~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
black<-aov(MEblack~treatment,data=MEs)
summary(black)
plot(MEblack~treatment,data=MEs)
```
```{r fig.height=3.5,fig.width=5}
  magenta<-aov(MEmagenta~treatment,data=MEs)
summary(magenta)
plot(MEmagenta~treatment,data=MEs)
```
```{r}
pink<-aov(MEpink~treatment,data=MEs)
summary(pink)
```
```{r}
red<-aov(MEred~treatment,data=MEs)
summary(red)
```