Rat E2 Project
========================================================
Hormones
-----------------------------------------------

```{r}
library(xtable)
arc<-read.csv("ARC TLDA for stats 7-20-13.csv") # load the data, this is the new data although I didn't change the date on the filename
which(colnames(arc)=="E2.RIA") #start of hormone data
which(colnames(arc)=="ACTH") #end
hormone<-arc[1:82,c(1,4:8,74:81)]
hor.table<-xtable(hormone)
#print(hor.table,type="html")
```
<!-- html table generated in R 2.15.2 by xtable 1.7-1 package -->
<!-- Tue Aug 27 10:33:21 2013 -->
<TABLE border=1>
<TR> <TH>  </TH> <TH> Sample.Name </TH> <TH> Group.code </TH> <TH> Group.ID </TH> <TH> Treatment..v.VEH..e.E2. </TH> <TH> age..y.young..m.MA..o.old. </TH> <TH> duration..3.3mo..6.6mo. </TH> <TH> E2.RIA </TH> <TH> LH </TH> <TH> FSH </TH> <TH> GH </TH> <TH> TSH </TH> <TH> PRL </TH> <TH> BDNF </TH> <TH> ACTH </TH>  </TR>
  <TR> <TD align="right"> 1 </TD> <TD> ABV19 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 23.41 </TD> <TD align="right"> 54.47 </TD> <TD align="right"> 45.16 </TD> <TD align="right"> 8922.33 </TD> <TD align="right"> 8.49 </TD> <TD align="right"> 7.23 </TD> <TD align="right"> 3.54 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 2 </TD> <TD> ABV20 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 23.93 </TD> <TD align="right"> 43.96 </TD> <TD align="right"> 54.35 </TD> <TD align="right"> 3188.26 </TD> <TD align="right"> 6.83 </TD> <TD align="right"> 0.99 </TD> <TD align="right"> 4.31 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 3 </TD> <TD> ABV25 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right">  </TD> <TD align="right"> 47.78 </TD> <TD align="right"> 53.70 </TD> <TD align="right"> 5792.02 </TD> <TD align="right"> 7.67 </TD> <TD align="right"> 7.26 </TD> <TD align="right"> 3.75 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 4 </TD> <TD> ABV26 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 17.05 </TD> <TD align="right"> 123.06 </TD> <TD align="right"> 113.18 </TD> <TD align="right"> 18976.14 </TD> <TD align="right"> 6.85 </TD> <TD align="right"> 2.10 </TD> <TD align="right"> 4.64 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 5 </TD> <TD> ABV76 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 41.83 </TD> <TD align="right"> 31.82 </TD> <TD align="right"> 36.42 </TD> <TD align="right"> 123101.20 </TD> <TD align="right"> 11.82 </TD> <TD align="right"> 0.29 </TD> <TD align="right"> 6.20 </TD> <TD align="right"> 0.04 </TD> </TR>
  <TR> <TD align="right"> 6 </TD> <TD> ABV85 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 39.33 </TD> <TD align="right"> 41.45 </TD> <TD align="right"> 38.02 </TD> <TD align="right"> 2239.19 </TD> <TD align="right"> 9.37 </TD> <TD align="right"> 1.20 </TD> <TD align="right"> 3.78 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 7 </TD> <TD> ABV86 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 29.47 </TD> <TD align="right"> 89.07 </TD> <TD align="right"> 175.03 </TD> <TD align="right"> 1928.57 </TD> <TD align="right"> 13.43 </TD> <TD align="right"> 11.85 </TD> <TD align="right"> 5.92 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 8 </TD> <TD> ABV87 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 21.84 </TD> <TD align="right"> 62.33 </TD> <TD align="right"> 89.96 </TD> <TD align="right"> 1544.47 </TD> <TD align="right"> 8.90 </TD> <TD align="right"> 1.20 </TD> <TD align="right"> 7.15 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 9 </TD> <TD> ABV88 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 7.15 </TD> <TD align="right"> 45.92 </TD> <TD align="right"> 42.81 </TD> <TD align="right"> 35317.26 </TD> <TD align="right"> 12.05 </TD> <TD align="right"> 0.54 </TD> <TD align="right"> 4.04 </TD> <TD align="right"> 0.06 </TD> </TR>
  <TR> <TD align="right"> 10 </TD> <TD> ABV89 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 19.28 </TD> <TD align="right"> 41.73 </TD> <TD align="right"> 47.73 </TD> <TD align="right"> 12808.42 </TD> <TD align="right"> 10.04 </TD> <TD align="right"> 1.02 </TD> <TD align="right"> 4.12 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 11 </TD> <TD> ABV90 </TD> <TD> YV3 </TD> <TD align="right">   1 </TD> <TD> v </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 15.13 </TD> <TD align="right"> 53.31 </TD> <TD align="right"> 495.72 </TD> <TD align="right"> 172082.03 </TD> <TD align="right"> 10.01 </TD> <TD align="right"> 0.64 </TD> <TD align="right"> 5.02 </TD> <TD align="right"> 0.04 </TD> </TR>
  <TR> <TD align="right"> 12 </TD> <TD> ABV17 </TD> <TD> YE3 </TD> <TD align="right">   2 </TD> <TD> e </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 140.51 </TD> <TD align="right"> 1.29 </TD> <TD align="right"> 23.96 </TD> <TD align="right"> 9047.69 </TD> <TD align="right"> 7.33 </TD> <TD align="right"> 80.53 </TD> <TD align="right"> 4.51 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 13 </TD> <TD> ABV18 </TD> <TD> YE3 </TD> <TD align="right">   2 </TD> <TD> e </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 101.44 </TD> <TD align="right"> 0.95 </TD> <TD align="right"> 16.36 </TD> <TD align="right"> 25366.94 </TD> <TD align="right"> 12.06 </TD> <TD align="right"> 64.86 </TD> <TD align="right"> 3.55 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 14 </TD> <TD> ABV23 </TD> <TD> YE3 </TD> <TD align="right">   2 </TD> <TD> e </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right">  </TD> <TD align="right"> 1.76 </TD> <TD align="right"> 16.72 </TD> <TD align="right"> 31932.36 </TD> <TD align="right"> 9.30 </TD> <TD align="right">  </TD> <TD align="right"> 1.92 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 15 </TD> <TD> ABV24 </TD> <TD> YE3 </TD> <TD align="right">   2 </TD> <TD> e </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right">  </TD> <TD align="right"> 14.62 </TD> <TD align="right"> 33.33 </TD> <TD align="right"> 50000.00 </TD> <TD align="right"> 19.76 </TD> <TD align="right">  </TD> <TD align="right"> 2.47 </TD> <TD align="right"> 0.04 </TD> </TR>
  <TR> <TD align="right"> 16 </TD> <TD> ABV71 </TD> <TD> YE3 </TD> <TD align="right">   2 </TD> <TD> e </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 96.57 </TD> <TD align="right"> 0.49 </TD> <TD align="right"> 11.73 </TD> <TD align="right"> 7996.20 </TD> <TD align="right"> 4.26 </TD> <TD align="right"> 54.81 </TD> <TD align="right"> 1.84 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 17 </TD> <TD> ABV72 </TD> <TD> YE3 </TD> <TD align="right">   2 </TD> <TD> e </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 80.44 </TD> <TD align="right"> 0.35 </TD> <TD align="right"> 10.96 </TD> <TD align="right"> 13546.25 </TD> <TD align="right"> 4.75 </TD> <TD align="right"> 47.01 </TD> <TD align="right"> 2.87 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 18 </TD> <TD> ABV75 </TD> <TD> YE3 </TD> <TD align="right">   2 </TD> <TD> e </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 133.97 </TD> <TD align="right"> 0.48 </TD> <TD align="right"> 23.75 </TD> <TD align="right"> 18818.18 </TD> <TD align="right"> 4.15 </TD> <TD align="right"> 6.70 </TD> <TD align="right"> 3.21 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 19 </TD> <TD> ABV79 </TD> <TD> YE3 </TD> <TD align="right">   2 </TD> <TD> e </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right"> 76.10 </TD> <TD align="right"> 1.17 </TD> <TD align="right"> 17.53 </TD> <TD align="right"> 12396.71 </TD> <TD align="right"> 2.01 </TD> <TD align="right"> 23.53 </TD> <TD align="right"> 4.37 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 20 </TD> <TD> ABV80 </TD> <TD> YE3 </TD> <TD align="right">   2 </TD> <TD> e </TD> <TD> y </TD> <TD align="right">  </TD> <TD align="right">  </TD> <TD align="right"> 0.45 </TD> <TD align="right"> 19.40 </TD> <TD align="right"> 7720.68 </TD> <TD align="right"> 4.89 </TD> <TD align="right"> 72.53 </TD> <TD align="right"> 3.84 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 21 </TD> <TD> ABV31 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 19.54 </TD> <TD align="right"> 37.39 </TD> <TD align="right"> 33.29 </TD> <TD align="right"> 2452.61 </TD> <TD align="right"> 4.77 </TD> <TD align="right"> 3.29 </TD> <TD align="right"> 3.58 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 22 </TD> <TD> ABV32 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 21.38 </TD> <TD align="right"> 28.72 </TD> <TD align="right"> 40.54 </TD> <TD align="right"> 4525.69 </TD> <TD align="right"> 5.45 </TD> <TD align="right"> 3.13 </TD> <TD align="right"> 3.26 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 23 </TD> <TD> ABV33 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 16.91 </TD> <TD align="right"> 18.23 </TD> <TD align="right"> 24.71 </TD> <TD align="right"> 1125.31 </TD> <TD align="right"> 5.78 </TD> <TD align="right"> 2.37 </TD> <TD align="right"> 2.93 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 24 </TD> <TD> ABV34 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 18.58 </TD> <TD align="right"> 383.43 </TD> <TD align="right"> 122.29 </TD> <TD align="right"> 50000.00 </TD> <TD align="right">  </TD> <TD align="right">  </TD> <TD align="right"> 2.31 </TD> <TD align="right"> 1.71 </TD> </TR>
  <TR> <TD align="right"> 25 </TD> <TD> ABV67 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 17.68 </TD> <TD align="right"> 19.67 </TD> <TD align="right"> 37.04 </TD> <TD align="right"> 4436.54 </TD> <TD align="right"> 4.30 </TD> <TD align="right"> 0.87 </TD> <TD align="right"> 2.09 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 26 </TD> <TD> ABV68 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 20.95 </TD> <TD align="right"> 100.79 </TD> <TD align="right"> 53.14 </TD> <TD align="right"> 50000.00 </TD> <TD align="right"> 42.22 </TD> <TD align="right"> 15.03 </TD> <TD align="right"> 3.64 </TD> <TD align="right"> 0.34 </TD> </TR>
  <TR> <TD align="right"> 27 </TD> <TD> ABV77 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 16.44 </TD> <TD align="right"> 70.02 </TD> <TD align="right"> 55.74 </TD> <TD align="right"> 50000.00 </TD> <TD align="right"> 17.38 </TD> <TD align="right"> 5.27 </TD> <TD align="right"> 4.05 </TD> <TD align="right"> 0.12 </TD> </TR>
  <TR> <TD align="right"> 28 </TD> <TD> ABV78 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 23.56 </TD> <TD align="right"> 65.15 </TD> <TD align="right"> 48.08 </TD> <TD align="right"> 50000.00 </TD> <TD align="right"> 15.05 </TD> <TD align="right"> 18.39 </TD> <TD align="right"> 3.64 </TD> <TD align="right"> 0.17 </TD> </TR>
  <TR> <TD align="right"> 29 </TD> <TD> ABV83 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right">  </TD> <TD align="right"> 66.55 </TD> <TD align="right"> 95.23 </TD> <TD align="right"> 50000.00 </TD> <TD align="right"> 13.60 </TD> <TD align="right"> 5.87 </TD> <TD align="right"> 3.74 </TD> <TD align="right"> 0.12 </TD> </TR>
  <TR> <TD align="right"> 30 </TD> <TD> ABV84 </TD> <TD> MV3 </TD> <TD align="right">   3 </TD> <TD> v </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 23.89 </TD> <TD align="right"> 33.97 </TD> <TD align="right"> 39.89 </TD> <TD align="right"> 4918.37 </TD> <TD align="right"> 9.60 </TD> <TD align="right"> 2.83 </TD> <TD align="right"> 5.02 </TD> <TD align="right"> 0.04 </TD> </TR>
  <TR> <TD align="right"> 31 </TD> <TD> ABV21 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 59.68 </TD> <TD align="right"> 1.66 </TD> <TD align="right"> 12.50 </TD> <TD align="right"> 15880.57 </TD> <TD align="right"> 2.89 </TD> <TD align="right"> 74.76 </TD> <TD align="right"> 1.94 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 32 </TD> <TD> ABV22 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 56.55 </TD> <TD align="right"> 1.27 </TD> <TD align="right"> 16.36 </TD> <TD align="right"> 6435.64 </TD> <TD align="right"> 3.13 </TD> <TD align="right"> 34.95 </TD> <TD align="right"> 2.44 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 33 </TD> <TD> ABV27 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 52.20 </TD> <TD align="right"> 8.54 </TD> <TD align="right"> 18.61 </TD> <TD align="right"> 25267.88 </TD> <TD align="right"> 5.49 </TD> <TD align="right"> 92.35 </TD> <TD align="right"> 2.39 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 34 </TD> <TD> ABV69 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right">  </TD> <TD align="right"> 0.75 </TD> <TD align="right"> 14.13 </TD> <TD align="right"> 4968.44 </TD> <TD align="right"> 7.71 </TD> <TD align="right"> 76.80 </TD> <TD align="right"> 2.67 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 35 </TD> <TD> ABV70 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 91.83 </TD> <TD align="right"> 1.10 </TD> <TD align="right"> 16.50 </TD> <TD align="right"> 19712.21 </TD> <TD align="right"> 2.28 </TD> <TD align="right"> 97.89 </TD> <TD align="right"> 5.62 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 36 </TD> <TD> ABV73 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 82.76 </TD> <TD align="right"> 0.50 </TD> <TD align="right"> 15.65 </TD> <TD align="right"> 12338.28 </TD> <TD align="right"> 4.74 </TD> <TD align="right">  </TD> <TD align="right"> 4.44 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 37 </TD> <TD> ABV74 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 36.97 </TD> <TD align="right"> 4.99 </TD> <TD align="right"> 24.30 </TD> <TD align="right"> 10556.52 </TD> <TD align="right"> 3.97 </TD> <TD align="right"> 3.22 </TD> <TD align="right"> 2.67 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 38 </TD> <TD> ABV81 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 78.69 </TD> <TD align="right"> 0.45 </TD> <TD align="right"> 13.27 </TD> <TD align="right"> 4983.46 </TD> <TD align="right"> 2.56 </TD> <TD align="right"> 55.69 </TD> <TD align="right"> 2.82 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 39 </TD> <TD> ABV82 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 74.34 </TD> <TD align="right"> 1.17 </TD> <TD align="right"> 14.61 </TD> <TD align="right"> 15454.46 </TD> <TD align="right"> 5.14 </TD> <TD align="right"> 99.40 </TD> <TD align="right"> 5.13 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 40 </TD> <TD> ABV91 </TD> <TD> ME3 </TD> <TD align="right">   4 </TD> <TD> e </TD> <TD> m </TD> <TD align="right">  </TD> <TD align="right"> 36.11 </TD> <TD align="right"> 0.50 </TD> <TD align="right"> 14.61 </TD> <TD align="right"> 1596.79 </TD> <TD align="right"> 1.36 </TD> <TD align="right"> 34.33 </TD> <TD align="right"> 2.14 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 41 </TD> <TD> ABV13 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 22.51 </TD> <TD align="right"> 27.12 </TD> <TD align="right"> 240.09 </TD> <TD align="right"> 2262.56 </TD> <TD align="right"> 7.07 </TD> <TD align="right"> 2.38 </TD> <TD align="right"> 2.53 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 42 </TD> <TD> ABV14 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 13.97 </TD> <TD align="right"> 26.52 </TD> <TD align="right"> 41.41 </TD> <TD align="right"> 9729.12 </TD> <TD align="right"> 5.63 </TD> <TD align="right"> 1.68 </TD> <TD align="right"> 5.41 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 43 </TD> <TD> ABV15 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 15.27 </TD> <TD align="right"> 21.40 </TD> <TD align="right"> 31.37 </TD> <TD align="right"> 739.24 </TD> <TD align="right"> 5.72 </TD> <TD align="right"> 3.80 </TD> <TD align="right"> 2.22 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 44 </TD> <TD> ABV16 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 18.26 </TD> <TD align="right"> 25.93 </TD> <TD align="right"> 29.94 </TD> <TD align="right"> 4539.72 </TD> <TD align="right"> 4.19 </TD> <TD align="right"> 2.47 </TD> <TD align="right"> 3.29 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 45 </TD> <TD> ABV49 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 17.52 </TD> <TD align="right"> 47.53 </TD> <TD align="right"> 71.55 </TD> <TD align="right"> 1248.68 </TD> <TD align="right"> 8.15 </TD> <TD align="right"> 6.47 </TD> <TD align="right"> 5.82 </TD> <TD align="right"> 0.04 </TD> </TR>
  <TR> <TD align="right"> 46 </TD> <TD> ABV61 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 29.08 </TD> <TD align="right"> 3.82 </TD> <TD align="right"> 63.68 </TD> <TD align="right"> 7544.22 </TD> <TD align="right"> 12.44 </TD> <TD align="right"> 3.35 </TD> <TD align="right"> 5.26 </TD> <TD align="right"> 0.05 </TD> </TR>
  <TR> <TD align="right"> 47 </TD> <TD> ABV62 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 14.41 </TD> <TD align="right"> 26.54 </TD> <TD align="right"> 39.09 </TD> <TD align="right"> 5463.62 </TD> <TD align="right"> 9.73 </TD> <TD align="right"> 1.71 </TD> <TD align="right"> 2.86 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 48 </TD> <TD> ABV63 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 11.55 </TD> <TD align="right"> 54.82 </TD> <TD align="right"> 105.64 </TD> <TD align="right"> 10045.29 </TD> <TD align="right"> 20.23 </TD> <TD align="right"> 5.00 </TD> <TD align="right"> 9.78 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 49 </TD> <TD> ABV64 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 16.64 </TD> <TD align="right"> 26.12 </TD> <TD align="right"> 43.95 </TD> <TD align="right"> 1172.31 </TD> <TD align="right"> 6.30 </TD> <TD align="right"> 1.68 </TD> <TD align="right"> 3.76 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 50 </TD> <TD> ABV65 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 9.89 </TD> <TD align="right"> 17.09 </TD> <TD align="right"> 43.16 </TD> <TD align="right"> 3225.23 </TD> <TD align="right"> 8.38 </TD> <TD align="right"> 2.98 </TD> <TD align="right"> 3.91 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 51 </TD> <TD> ABV66 </TD> <TD> MV6 </TD> <TD align="right">   5 </TD> <TD> v </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 20.88 </TD> <TD align="right"> 18.66 </TD> <TD align="right"> 23.24 </TD> <TD align="right"> 2385.76 </TD> <TD align="right"> 4.21 </TD> <TD align="right"> 5.39 </TD> <TD align="right"> 3.51 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 52 </TD> <TD> ABV09 </TD> <TD> ME6 </TD> <TD align="right">   6 </TD> <TD> e </TD> <TD> o </TD> <TD align="right">   6 </TD> <TD align="right"> 55.89 </TD> <TD align="right"> 1.03 </TD> <TD align="right"> 17.10 </TD> <TD align="right"> 4025.75 </TD> <TD align="right"> 1.48 </TD> <TD align="right"> 14.06 </TD> <TD align="right"> 2.11 </TD> <TD align="right"> 0.00 </TD> </TR>
  <TR> <TD align="right"> 53 </TD> <TD> ABV10 </TD> <TD> ME6 </TD> <TD align="right">   6 </TD> <TD> e </TD> <TD> o </TD> <TD align="right">   6 </TD> <TD align="right"> 64.16 </TD> <TD align="right"> 1.32 </TD> <TD align="right"> 17.38 </TD> <TD align="right"> 4533.35 </TD> <TD align="right"> 2.72 </TD> <TD align="right"> 41.97 </TD> <TD align="right"> 2.80 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 54 </TD> <TD> ABV11 </TD> <TD> ME6 </TD> <TD align="right">   6 </TD> <TD> e </TD> <TD> o </TD> <TD align="right">   6 </TD> <TD align="right"> 60.86 </TD> <TD align="right"> 2.76 </TD> <TD align="right"> 14.29 </TD> <TD align="right"> 25823.87 </TD> <TD align="right"> 3.24 </TD> <TD align="right"> 41.24 </TD> <TD align="right"> 2.42 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 55 </TD> <TD> ABV12 </TD> <TD> ME6 </TD> <TD align="right">   6 </TD> <TD> e </TD> <TD> o </TD> <TD align="right">   6 </TD> <TD align="right"> 43.06 </TD> <TD align="right"> 2.58 </TD> <TD align="right"> 13.91 </TD> <TD align="right"> 11758.89 </TD> <TD align="right"> 3.44 </TD> <TD align="right"> 61.67 </TD> <TD align="right"> 2.33 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 56 </TD> <TD> ABV50 </TD> <TD> ME6 </TD> <TD align="right">   6 </TD> <TD> e </TD> <TD> o </TD> <TD align="right">   6 </TD> <TD align="right"> 70.92 </TD> <TD align="right"> 14.51 </TD> <TD align="right"> 12.12 </TD> <TD align="right"> 50000.00 </TD> <TD align="right"> 14.42 </TD> <TD align="right">  </TD> <TD align="right"> 2.19 </TD> <TD align="right"> 0.38 </TD> </TR>
  <TR> <TD align="right"> 57 </TD> <TD> ABV51 </TD> <TD> ME6 </TD> <TD align="right">   6 </TD> <TD> e </TD> <TD> o </TD> <TD align="right">   6 </TD> <TD align="right">  </TD> <TD align="right"> 0.16 </TD> <TD align="right"> 9.50 </TD> <TD align="right"> 6325.06 </TD> <TD align="right"> 2.30 </TD> <TD align="right"> 28.41 </TD> <TD align="right"> 1.43 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 58 </TD> <TD> ABV52 </TD> <TD> ME6 </TD> <TD align="right">   6 </TD> <TD> e </TD> <TD> o </TD> <TD align="right">   6 </TD> <TD align="right"> 76.21 </TD> <TD align="right"> 0.32 </TD> <TD align="right"> 14.41 </TD> <TD align="right"> 22004.19 </TD> <TD align="right"> 1.56 </TD> <TD align="right"> 71.41 </TD> <TD align="right"> 3.23 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 59 </TD> <TD> ABV53 </TD> <TD> ME6 </TD> <TD align="right">   6 </TD> <TD> e </TD> <TD> o </TD> <TD align="right">   6 </TD> <TD align="right"> 47.35 </TD> <TD align="right"> 4.31 </TD> <TD align="right"> 16.22 </TD> <TD align="right"> 22262.22 </TD> <TD align="right"> 4.94 </TD> <TD align="right"> 89.10 </TD> <TD align="right"> 3.69 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 60 </TD> <TD> ABV54 </TD> <TD> ME6 </TD> <TD align="right">   6 </TD> <TD> e </TD> <TD> o </TD> <TD align="right">   6 </TD> <TD align="right"> 82.33 </TD> <TD align="right"> 1.92 </TD> <TD align="right"> 18.56 </TD> <TD align="right"> 16298.70 </TD> <TD align="right"> 5.45 </TD> <TD align="right">  </TD> <TD align="right"> 3.71 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 61 </TD> <TD> ABV01 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right">  </TD> <TD align="right"> 0.69 </TD> <TD align="right"> 6.93 </TD> <TD align="right"> 570.52 </TD> <TD align="right"> 0.07 </TD> <TD align="right"> 26.44 </TD> <TD align="right"> 0.97 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 62 </TD> <TD> ABV02 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 74.60 </TD> <TD align="right"> 0.13 </TD> <TD align="right"> 4.93 </TD> <TD align="right"> 14084.66 </TD> <TD align="right"> 1.08 </TD> <TD align="right"> 22.51 </TD> <TD align="right"> 3.02 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 63 </TD> <TD> ABV03 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 44.46 </TD> <TD align="right"> 0.52 </TD> <TD align="right"> 14.94 </TD> <TD align="right"> 4307.62 </TD> <TD align="right"> 2.57 </TD> <TD align="right"> 31.22 </TD> <TD align="right"> 1.67 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 64 </TD> <TD> ABV04 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 43.47 </TD> <TD align="right"> 1.33 </TD> <TD align="right"> 13.35 </TD> <TD align="right"> 8808.48 </TD> <TD align="right"> 2.72 </TD> <TD align="right"> 53.91 </TD> <TD align="right"> 3.36 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 65 </TD> <TD> ABV41 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 52.83 </TD> <TD align="right"> 0.52 </TD> <TD align="right"> 6.44 </TD> <TD align="right"> 50000.00 </TD> <TD align="right"> 7.52 </TD> <TD align="right"> 76.93 </TD> <TD align="right"> 2.72 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 66 </TD> <TD> ABV42 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 65.02 </TD> <TD align="right"> 0.24 </TD> <TD align="right"> 6.06 </TD> <TD align="right"> 17605.77 </TD> <TD align="right"> 1.42 </TD> <TD align="right"> 26.29 </TD> <TD align="right"> 2.86 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 67 </TD> <TD> ABV43 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 34.21 </TD> <TD align="right"> 0.54 </TD> <TD align="right"> 8.07 </TD> <TD align="right"> 5643.66 </TD> <TD align="right"> 2.16 </TD> <TD align="right"> 31.05 </TD> <TD align="right"> 1.64 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 68 </TD> <TD> ABV44 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 73.21 </TD> <TD align="right"> 0.92 </TD> <TD align="right"> 15.13 </TD> <TD align="right"> 9394.77 </TD> <TD align="right"> 2.00 </TD> <TD align="right"> 21.17 </TD> <TD align="right"> 5.07 </TD> <TD align="right"> 0.05 </TD> </TR>
  <TR> <TD align="right"> 69 </TD> <TD> ABV45 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 107.07 </TD> <TD align="right"> 0.24 </TD> <TD align="right"> 5.96 </TD> <TD align="right"> 10349.03 </TD> <TD align="right"> 2.22 </TD> <TD align="right"> 76.44 </TD> <TD align="right"> 3.00 </TD> <TD align="right"> 0.04 </TD> </TR>
  <TR> <TD align="right"> 70 </TD> <TD> ABV46 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 65.66 </TD> <TD align="right"> 0.20 </TD> <TD align="right"> 10.54 </TD> <TD align="right"> 5956.27 </TD> <TD align="right"> 1.39 </TD> <TD align="right"> 62.38 </TD> <TD align="right"> 2.88 </TD> <TD align="right"> 0.04 </TD> </TR>
  <TR> <TD align="right"> 71 </TD> <TD> ABV47 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 72.10 </TD> <TD align="right"> 1.12 </TD> <TD align="right"> 15.84 </TD> <TD align="right"> 4488.20 </TD> <TD align="right"> 2.01 </TD> <TD align="right"> 56.35 </TD> <TD align="right"> 2.55 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 72 </TD> <TD> ABV48 </TD> <TD> MV3E3 </TD> <TD align="right">   7 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">  </TD> <TD align="right"> 68.67 </TD> <TD align="right"> 0.17 </TD> <TD align="right"> 8.42 </TD> <TD align="right"> 8771.18 </TD> <TD align="right"> 3.20 </TD> <TD align="right"> 72.62 </TD> <TD align="right"> 1.83 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 73 </TD> <TD> ABV05 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 21.46 </TD> <TD align="right"> 27.38 </TD> <TD align="right"> 31.08 </TD> <TD align="right"> 10462.39 </TD> <TD align="right"> 5.21 </TD> <TD align="right"> 8.47 </TD> <TD align="right"> 2.62 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 74 </TD> <TD> ABV06 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 17.84 </TD> <TD align="right"> 11.85 </TD> <TD align="right"> 11.80 </TD> <TD align="right"> 5936.84 </TD> <TD align="right"> 5.49 </TD> <TD align="right"> 9.54 </TD> <TD align="right"> 2.13 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 75 </TD> <TD> ABV07 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 15.98 </TD> <TD align="right"> 16.78 </TD> <TD align="right"> 27.27 </TD> <TD align="right"> 1416.68 </TD> <TD align="right"> 4.97 </TD> <TD align="right"> 2.46 </TD> <TD align="right"> 2.20 </TD> <TD align="right"> 0.02 </TD> </TR>
  <TR> <TD align="right"> 76 </TD> <TD> ABV08 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 18.76 </TD> <TD align="right"> 16.10 </TD> <TD align="right"> 16.68 </TD> <TD align="right"> 3043.80 </TD> <TD align="right"> 6.24 </TD> <TD align="right"> 3.92 </TD> <TD align="right"> 2.73 </TD> <TD align="right"> 0.03 </TD> </TR>
  <TR> <TD align="right"> 77 </TD> <TD> ABV55 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 22.78 </TD> <TD align="right"> 25.78 </TD> <TD align="right"> 45.80 </TD> <TD align="right"> 3238.77 </TD> <TD align="right"> 10.27 </TD> <TD align="right"> 13.99 </TD> <TD align="right"> 3.00 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 78 </TD> <TD> ABV56 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 24.31 </TD> <TD align="right"> 29.97 </TD> <TD align="right"> 88.45 </TD> <TD align="right"> 7227.28 </TD> <TD align="right"> 10.00 </TD> <TD align="right"> 2.36 </TD> <TD align="right"> 4.56 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 79 </TD> <TD> ABV57 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 32.46 </TD> <TD align="right"> 30.89 </TD> <TD align="right"> 38.91 </TD> <TD align="right"> 7944.46 </TD> <TD align="right"> 11.85 </TD> <TD align="right"> 7.06 </TD> <TD align="right"> 4.57 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 80 </TD> <TD> ABV58 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 19.03 </TD> <TD align="right"> 34.15 </TD> <TD align="right"> 41.92 </TD> <TD align="right"> 1819.70 </TD> <TD align="right"> 10.79 </TD> <TD align="right"> 9.28 </TD> <TD align="right"> 3.55 </TD> <TD align="right"> 0.01 </TD> </TR>
  <TR> <TD align="right"> 81 </TD> <TD> ABV59 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 12.37 </TD> <TD align="right"> 34.17 </TD> <TD align="right"> 45.63 </TD> <TD align="right"> 7122.87 </TD> <TD align="right"> 9.12 </TD> <TD align="right"> 5.34 </TD> <TD align="right"> 7.64 </TD> <TD align="right">  </TD> </TR>
  <TR> <TD align="right"> 82 </TD> <TD> ABV60 </TD> <TD> ME3V3 </TD> <TD align="right">   8 </TD> <TD>  </TD> <TD> o </TD> <TD align="right">   3 </TD> <TD align="right"> 12.64 </TD> <TD align="right"> 9.87 </TD> <TD align="right"> 58.51 </TD> <TD align="right"> 7419.53 </TD> <TD align="right"> 11.69 </TD> <TD align="right"> 7.70 </TD> <TD align="right"> 7.21 </TD> <TD align="right"> 0.02 </TD> </TR>
   </TABLE>

## Grouping
I'm going to switch how some of the grouping variables are encoded. For the E2 effect I'm going to include all the animals even the ones that switch from E2 to VEH (and vice-versa) using the drug that they had most recently as their group because that is the predominant signal I'm seeing in the data. For the duration effect I'm going to include animals of all ages but I'm going to exclude the switching treatment for now. For the age effect I'm only going to use two levels "young"" and "mature" because the difference between "mature" and "old" is captured by the duration variable.

code to change the groupings:
```{r}
names(hormone)[4:6]<-c("e2","age","duration")
hormone$e2[61:82]<-as.factor(ifelse(hormone$Group.ID[61:82]==7,yes="e",no="v"))
hormone$age[which(hormone$age=="o")]<-factor("m")
hormone$duration<-factor(c(rep(3,length(1:40)),rep(6,length(41:60)),rep(NA,length(61:82))))
```
## Comparisons
I think it makes sense to plan the comparisons here rather than doing all pairwise comparisons because some of them don't necessarily make sense to compare. I'm thinking there should be two models, one where we look at E2, age and duration as seperate and potentially interacting factors. The second model will partion them into their eight treatment groups and look at the following comparisons: 

* YV3 (1) vs YE3 (2)
* MV3 (3) vs ME3 (4)
* MV6 (5) vs ME6 (6)
* MV3E3 (7) vs ME3V3 (8)

## Normality
None of the hormone data are normally distributed (or multivariate normal). Log transformation improves the normality but not completly. This leaves us with a few options, we can run the MANOVA/ANOVAs on the log transformed data and hope that our results are robust to that, we can use a non-parametric kruskal-wallis type test or we can use permutation.

Normality tests on the transformed data (either sqrt or log depending on what gave the best results):
```{r}
attach(hormone)
shapiro.test(log10(E2.RIA))
shapiro.test(log10(LH))
shapiro.test(log10(FSH))
shapiro.test(log10(GH))
shapiro.test(sqrt(TSH))
shapiro.test(log10(PRL))
shapiro.test(log10(BDNF))
shapiro.test(log10(ACTH))

thormone<-hormone
thormone[,7:14]<-as.data.frame(cbind(log10(E2.RIA),log10(LH),log10(FSH),log10(GH),sqrt(TSH),log10(PRL),log10(BDNF),log10(ACTH)))
detach(hormone)
```
multivariate test of normality... cannot have missing data so here I'm going to impute the missing data as an average of the three nearest neighbors:

```{r}
library(mvnormtest)
library(imputation)
imputed<-kNNImpute(thormone[,7:14],3)$x
mshapiro.test(t(as.matrix((imputed)))) # not normal or multivariate normal
```
## parametric anovas on log/sqrt transformed data ignoring the normality problem for now:
Taking out the treatment-switching for now...
```{r}
thormone_with_switch<-thormone
thormone<-thormone[thormone$Group.ID<7,]
#first model looking E2, age and duration as well as the interactions
man1<-manova(cbind(E2.RIA,LH,FSH,GH,TSH,PRL,BDNF,ACTH)~age*e2*duration,data=thormone)#not normal
summary(man1,test="Wilks",intercept=TRUE)
summary.aov(man1)
```

```{r}
#second model looking E2 and age as well as the interaction
man2<-manova(cbind(E2.RIA,LH,FSH,GH,TSH,PRL,BDNF,ACTH)~age*e2,data=thormone)#not normal
summary(man2,test="Wilks",intercept=TRUE)
summary.aov(man2)
aov1<-aov(E2.RIA~age*duration*e2,data=thormone)
summary(lm(E2.RIA~age*duration*e2,data=thormone))
```

the design isn't fully crossed... duration effect is a bit of a problem because there is no duration-6mo x young group. This makes it impossible to estimate the 3 way interaction.

## just looking at E2 + Age for now
```{r fig.width=7, fig.height=6}
thormone$age<-as.factor(as.character(thormone$age))
boxplot(E2.RIA~age*e2,data=thormone,ylab="E2.RIA (log)",main="without switching group")
```
```{r fig.width=7, fig.height=6}
thormone_with_switch$age<-as.factor(as.character(thormone_with_switch$age))
boxplot(E2.RIA~age*e2,data=thormone_with_switch,ylab="E2.RIA (log)",main="with switching group")
```

```{r fig.width=7, fig.height=6}
boxplot(LH~age*e2,data=thormone,ylab="LH (log)",main="without switching group")
```
```{r fig.width=7, fig.height=6}
boxplot(LH~age*e2,data=thormone_with_switch,ylab="LH (log)",main="with switching group")
```

```{r fig.width=7, fig.height=6}
boxplot(FSH~age*e2,data=thormone,ylab="FSH (log)",main="without switching group")
```
```{r fig.width=7, fig.height=6}
boxplot(FSH~age*e2,data=thormone_with_switch,ylab="FSH (log)",main="with switching group")
```

```{r fig.width=7, fig.height=6}
boxplot(GH~age*e2,data=thormone,ylab="GH",main="without switching group")
```
```{r fig.width=7, fig.height=6}
boxplot(GH~age*e2,data=thormone_with_switch,ylab="GH",main="with switching group")
```

```{r fig.width=7, fig.height=6}
boxplot(TSH~age*e2,data=thormone,ylab="TSH (sqrt)",main="without switching group")
```
```{r fig.width=7, fig.height=6}
boxplot(TSH~age*e2,data=thormone_with_switch,ylab="TSH (sqrt)",main="with switching group")
```


```{r fig.width=7, fig.height=6}
boxplot(PRL~age*e2,data=thormone,ylab="PRL (log10)",main="without switching group")
```
```{r fig.width=7, fig.height=6}
boxplot(PRL~age*e2,data=thormone_with_switch,ylab="PRL (log10)",main="with switching group")
```


```{r fig.width=7, fig.height=6}
boxplot(BDNF~age*e2,data=thormone,ylab="BDNF (log10)",main="without switching group")
```
```{r fig.width=7, fig.height=6}
boxplot(BDNF~age*e2,data=thormone_with_switch,ylab="BDNF (log10)",main="with switching group")
```


```{r fig.width=7, fig.height=6}
boxplot(ACTH~age*e2,data=thormone,ylab="ACTH (log10)",main="without switching group")
```
```{r fig.width=7, fig.height=6}
boxplot(ACTH~age*e2,data=thormone_with_switch,ylab="ACTH (log10)",main="with switching group")
```

## Addressing normality
I think the best option is to compare the results of each single AOV against a permutation based method. If we had only one factor we could use the kruskal wallis (could use it for the other model?) 

Note from lmPerm R package: "In those cases where the normal theory assumptions are adequately approximated, permutation tests are indistinguishable from the usual F-tests. Section 3.2 shows simulations illustrating this. In those cases where the p-values from permutation tests differ substantially from those for F-tests, a careful examination of the data is usually worthwhile."

```{r}
library(lmPerm)
#regular ANVOA:
anova(lm(E2.RIA~age*e2,data=thormone))
#permutation:
anova(lmp(E2.RIA~age*e2,data=thormone,perm="exact"))

#regular ANVOA:
anova(lm(LH~age*e2,data=thormone))
#permutation:
anova(lmp(LH~age*e2,data=thormone,perm="exact"))

#regular ANVOA:
anova(lm(FSH~age*e2,data=thormone))
#permutation:
anova(lmp(FSH~age*e2,data=thormone,perm="exact"))

#regular ANVOA:
anova(lm(GH~age*e2,data=thormone))
#permutation:
anova(lmp(GH~age*e2,data=thormone,perm="exact"))

#regular ANVOA:
anova(lm(TSH~age*e2,data=thormone))
#permutation:
anova(lmp(TSH~age*e2,data=thormone,perm="exact"))

#regular ANVOA:
anova(lm(PRL~age*e2,data=thormone))
#permutation:
anova(lmp(PRL~age*e2,data=thormone,perm="exact"))

#regular ANVOA:
anova(lm(BDNF~age*e2,data=thormone))
#permutation:
anova(lmp(BDNF~age*e2,data=thormone,perm="exact"))


#regular ANVOA:
anova(lm(ACTH~age*e2,data=thormone))
#permutation:
anova(lmp(ACTH~age*e2,data=thormone,perm="exact"))
```

## Okay now for the second type of model keeping all eight groups seperate and testing the paired comparisons

```{r}
library(multcomp)
thormone_with_switch$Group.code<-as.factor(as.character(thormone_with_switch$Group.code))
for(response in 7:14){
  regaov<-aov(thormone_with_switch[,response]~Group.code,data=thormone_with_switch)
  rname<-(names(thormone_with_switch)[response])
  print(rname)
  boxplot(thormone_with_switch[,response]~Group.code,data=thormone_with_switch,main=rname)
  print(summary(regaov))
  mc<-glht(regaov,linfct=mcp(Group.code = c("MV3 - ME3 = 0", "MV6 - ME6 = 0","ME3V3 - MV3E3 = 0","YV3 - YE3 = 0")))
  print(summary(mc))
}

```

```{r}
library(MASS)
scale_imputed<-as.data.frame(scale(imputed))
attach(scale_imputed)
fitJack<-lda(thormone_with_switch$Group.code~E2.RIA+LH+FSH+GH+TSH+PRL+BDNF+ACTH,na.action="na.omit",CV=TRUE)
ct <- table(thormone_with_switch$Group.code, fitJack$class)
fitJack

print(ct)
diag(prop.table(ct, 1))
# total percent correct
sum(diag(prop.table(ct)))

fit<-lda(thormone_with_switch$Group.code~E2.RIA+LH+FSH+GH+TSH+PRL+BDNF+ACTH,na.action="na.omit",CV=F)

plot(fit,dimen=2)
```

```{r fig.height=11,fig.width=8.5}
plot(fit,dimen=3)
```